### 2016-04-26 msmfs.tex
Added much material
Is MSMFS needed?
Equations about the Hessian etc
More recommendations

\documentclass[11pt,a4paper]{article}

@@

@@

\newcommand{\dg}{^\dag}
\newcommand{\X}{\vec{x}}
\newcommand{\Xd}{\vec{{x}^\dag}}
\newcommand{\B}{\vec{b}}
\newcommand{\Bd}{\vec{b^\dag}}
\newcommand{\V}{\vec{V}}
\newcommand{\Vd}{\vec{V^\dag}}
\newcommand{\A}{{\tens{A}{s}}}
\newcommand{\Ad}{{\tens{A^\dag}{s}}}
\newcommand{\F}{{\tens{F}{s}}}
\newcommand{\Fd}{{\tens{F^\dag}{s}}}
\newcommand{\He}{{\tens{H}{s}}}
\newcommand{\Sa}{{\tens{S}{s}}}
\newcommand{\Sd}{{\tens{S^\dag}{s}}}
\newcommand{\Sna}{\tens{{S_{\nu}}{s}}}
\newcommand{\Snd}{\tens{{S_{\nu}^\dag}{s}}}
\newcommand{\T}{{\tens{T}{s}}}
\newcommand{\W}{{\tens{W}{s}}}
\newcommand{\Wd}{{\tens{W^\dag}{s}}}
\newcommand{\Pb}{{\vec{P}}}

@@

% Add here the symbols used in your document
%\clearpage \sdplistofsymbols
%\begin{basedescript}{\desclabelstyle{\pushlabel}\desclabelwidth{6em}}
%\item[$\bytes$] Length of a complex number in bytes \vspace{-0.2cm}
%\item[$\cbytes$] Length of a complex number in bytes \vspace{-0.2cm}
%\item[$\rbytes$] Length of a real number in bytes \vspace{-0.2cm}
%\item[$\bl$] Baseline length / GPU memory bandwidth \vspace{-0.2cm}
%\item[$\maxbl$] Maximum baseline length \vspace{-0.2cm}
%\item[$\membw$] Memory bandwidth \vspace{-0.2cm}
%\item [$\pciebw$] PCI-e bandwidth \vspace{-0.2cm}
%\item[$\viscompr$] Compression factor of visibilities for grdding for facets
%  \textcolor{red}{This symbol is not used in the document, consider removing.}
%  \vspace{-0.2cm}
%\item[$\vism$] Matrix of observed visibilities \vspace{-0.2cm}
%\item[$\da$] Extent of aperture \vspace{-0.2cm}
%\item[$\ds$] Diameter of antennas/stations \vspace{-0.2cm}
%\item[$\reff$] Reference frequency\vspace{-0.2cm}
%\item[$\fcc$] Convolution kernel computational cost\vspace{-0.2cm}
%\item[$\fci$] Factor between bandwidth and FLOPS (bytes/operations)\vspace{-0.2cm}
%\item[$\fgrid$] Gridding cost\vspace{-0.2cm}
%\item[$\fpatch$] Fraction of the maximum baseline below which the $\uv$
%  coverage is almost filled\vspace{-0.2cm}
%\item[$\fpr$] Cost for phase rotation\vspace{-0.2cm}
%\item[$\gainm$] 2$\times$2 block diagonal matrix of complex
%  gains\vspace{-0.2cm}
%\item[$\modskyvism$] Matrix of model sky visibilities\vspace{-0.2cm}
%\item[$\skymodbufsize$] Size of the sky model buffer\vspace{-0.2cm}
%\item[$\visbufsize$] Size of the visibility buffer \vspace{-0.2cm}
%\item[$\visbufsizespec$] Size of the visibility buffer needed for spectral line processing\vspace{-0.2cm}
%\item[$\visbufsizecont$] Size of the visibility buffer needed for continuum imaging\vspace{-0.2cm}
%\item[$\bufsize$] Size of buffer \vspace{-0.2cm}
%\item[$\slowmemsize$] Size of slow working memory \vspace{-0.2cm}
%\item[$\cumemsize$] Size of working memory of compute unit \vspace{-0.2cm}
%\item[$\imgsize$] Size of image (bytes) \vspace{-0.2cm}
%\item[$\imgsizeoneax$] Image size along one axis \vspace{-0.2cm}
%\item[$\targgridsize$] Target grid size (per facet) \vspace{-0.2cm}
%\item[$\uvgridsize$] Size of the $\uv$ grid (bytes) \vspace{-0.2cm}
%\item[$\vissize$] Size of visibility data (bytes) \vspace{-0.2cm}
%\item[$\wgridsize$] Size of weight grid (bytes)\vspace{-0.2cm}
%\item[$\gridcachesize$] Cache size of gridding kernel \vspace{-0.2cm}
%\item[$\nbins$] Number of bins \vspace{-0.2cm}
%\item[$\ntel$] Number of antennas or stations in the array
%  \vspace{-0.2cm}
%\item[$\nateam$] Number of A-team sources \vspace{-0.2cm}
%\item[$\ngridpakern$] Number of grid points for A-kernel support (along one
%  axis) \vspace{-0.2cm}
%\item[$\naccvis$] Number of accesses per visibility \vspace{-0.2cm}
%\item[$\navg$] Number of averages\vspace{-0.2cm}
%\item[$\nbeam$] Number of beams simultaneously observed by the array
%  \vspace{-0.2cm}
%\item[$\nbits$] Number of bits\vspace{-0.2cm}
%\item[$\rma$] Number of flops in real multiply and add\vspace{-0.2cm}
%\item[$\cma$] Number of flops in complex multiply and add\vspace{-0.2cm}
%\item[$\nbl$] Number of baselines \vspace{-0.2cm}
%\item[$\nbyteacc$] Number of bytes per access \vspace{-0.2cm}
%\item[$\nbyte$] Number of bytes \vspace{-0.2cm}
%\item[$\nbyteperpix$] Number of bytes per pixel \vspace{-0.2cm}
%\item[$\nbytepervis$] Number of bytes per individual visibility 
%
%\item[$\nch$] Number of channels \vspace{-0.2cm}
%\item[$\ncompu$] Number of compute units \vspace{-0.2cm}
%\item[$\ndatait$] Number of data items required per source (e.g. position)\vspace{-0.2cm}
%\item[$\ngridp$] Number of grid points along one axis for the gridding
%  convolution function \vspace{-0.2cm}
%\item[$\nfch$] Number of frequency channels/sub-bands \vspace{-0.2cm}
%\item[$\nfchcorr$] Number of frequency channels output by the correlator
%  \vspace{-0.2cm}
%\item[$\nfchgrid$] Number of frequency channels to be gridded
%  \vspace{-0.2cm}
%\item[$\nfchdegrid$] Number of frequency channels to be de-gridded
%  \vspace{-0.2cm}
%\item[$\nfchout$] Number of frequency channels to be FFT-ed and reprojected
%  (output) or output from the CSP pulsar timing backend\vspace{-0.2cm}
%\item[$\nfchdist$] Minimum number of frequency channels to maintain distributability in the SDP
%\item[$\nfkernel$] Number of convolution kernels needed to cover frequency axis for a given baseline
%\item[$\nfacet$] Number of facets along one axis \vspace{-0.2cm}
%\item[$\nflop$] Number of floating point operations
%  \vspace{-0.2cm}
%\item[$\nflopsamp$] Number of floating point operations per sample
%  \vspace{-0.2cm}
%\item[$\nflopvis$] Number of floating point operations per visibility
%\item[$\npixgrid$] Number of pixels in the grid (i.e. the square of the linear
%  dimension) \vspace{-0.2cm}
%\item[$\ngridwkern$] Number of grid points for w-kernel support (along one
%  axis) \vspace{-0.2cm}
%\item[$\nit$] Number of iterations \vspace{-0.2cm}
%\item[$\nawkern$] Linear size of combined A and $\w$-kernels \vspace{-0.2cm}
%\item[$\nmaj$] Number of major CLEAN cycles (within each self-cal loop if appropriate) \vspace{-0.2cm}
%\item[$\nselfcal$] Number of self-calibration cycles \vspace{-0.2cm}
%\item[$\nmin$] Number of minor CLEAN cycles (within each major cycle loop if appropriate)\vspace{-0.2cm}
%\item[$\facgridoffdiagmuller$] Factor to account for gridding of off-diagonal
%  M\"uller terms \vspace{-0.2cm}
%\item[$\nops$] Number of operations \vspace{-0.2cm}
%\item[$\npatchpix$] Number of pixels of PSF to be used in minor cycle
%  decorrelation \vspace{-0.2cm}
%\item[$\npix$] Number of pixels along one axis of a grid\vspace{-0.2cm}
%\item[$\npixfacet$] Number of pixels on one side of a facet \vspace{-0.2cm}
%\item[$\npol$] Number of polarisation products \vspace{-0.2cm}
%\item[$\nsource$] Number of sources \vspace{-0.2cm}
%\item[$\ntslot$] Number of time slots \vspace{-0.2cm}
%\item[$\ntaylor$] Number of Taylor terms (in MS-MFS)\vspace{-0.2cm}
%\item[$\ntslotave$] Number of time slots that are averaged (in de-mixing)
%  \vspace{-0.2cm}
%\item[$\nvis$] Number of visibilities in an observation \vspace{-0.2cm}
%\item[$\qfacbw$] Quality factor \vspace{-0.2cm}
%\item[$\qfacreuse$] Factor to allow for the reuse of functions between
%  neighbouring frequency channels \vspace{-0.2cm}
%\item[$\qfacfov$] Quality factor defining how much bigger the diameter of the
%  field of view is than the first zero of the Airy function \vspace{-0.2cm}
%\item[$\facgridovers$] Gridding oversampling factor \vspace{-0.2cm}
%\item[$\facmsmf$] Factor to account for multiple subtractions (multi-frequency
%  multi-scale CLEAN\vspace{-0.2cm}
%\item[$\qfacbeamovers$] Quality factor defining the oversampling of the
%  synthesised beam \vspace{-0.2cm}
%\item[$\qfac$] Quality factor \vspace{-0.2cm}
%\item[$\Qkernel$] Convolution kernel quality factor, for kernel re-use in frequency
%\item[$\corrfacw$] Correction factor for w values (0 to 1) \vspace{-0.2cm}
%\item[$\earthrad$] Radius of the Earth ($6400\times 10^3$\,m) \vspace{-0.2cm}
%\item[$\covm$] Covariance matrix \vspace{-0.2cm}
%\item[$\covmfilt$] Filtered covariance matrix \vspace{-0.2cm}
%\item[$\comprate$] Compute rate (FLOPS) \vspace{-0.2cm}
%\item[$\compratespec$] Compute rate (FLOPS) for spectral line processing \vspace{-0.2cm}
%\item[$\compratecont$] Compute rate (FLOPS) for continuum imaging \vspace{-0.2cm}
%\item[$\compratefast$] Compute rate (FLOPS) for fast imaging \vspace{-0.2cm}
%\item[$\bwtoworkmem$] Bandwidth to working memory \vspace{-0.2cm}
%\item[$\maxmembw$] Maximum memory bandwidth of compute unit \vspace{-0.2cm}
%\item[$\peakflop$] Peak FLOPS of compute unit \vspace{-0.2cm}
%\item[$\maxiobwbuff$] Maximum I/O bandwidth of compute unit to buffer
%  \vspace{-0.2cm}
%\item[$\bwbuff$] Bandwidth to buffer \vspace{-0.2cm}
%\item[$\bwbuffcont$] Bandwidth to buffer for continuum imaging\vspace{-0.2cm}
%\item[$\bwbuffspec$] Bandwidth to buffer for spectral line processing\vspace{-0.2cm}
%\item[$\convkernflops$] Convolution kernel compute rate (FLOPS)\vspace{-0.2cm}
%\item[$\fftflops$] FFT compute rate (FLOPS)\vspace{-0.2cm}
%\item[$\griddingflops$] Gridding compute rate (FLOPS)\vspace{-0.2cm}
%\item[$\phaserotflops$] Phase rotation compute rate (FLOPS)\vspace{-0.2cm}
%\item[$\reprojflops$] Re-projection compute rate (FLOPS)\vspace{-0.2cm}
%\item[$\rcorr$] Correlator output rate\vspace{-0.2cm} TRY TO REMOVE
%\item[$\facetoverlap$]  ratio for linear size increase of facets to account for overlap, typically 1.2-1.5 \vspace{-0.2cm}
%\item[$\telsensit$] Telescope sensitivity \vspace{-0.2cm} TRY TO REMOVE
%\item[$\corrdumpt$] Correlator dump time (sec) \vspace{-0.2cm}
%\item[$\compt$] Computation time \vspace{-0.2cm}
%\item[$\gridkernupdate$] Update timescale for gridding kernel \vspace{-0.2cm}
%\item[$\lsmupdate$] Update timescale for the LSM \vspace{-0.2cm}
%\item[$\obst$] Observation time \vspace{-0.2cm}
%\item[$\snapshot$] Snapshot duration \vspace{-0.2cm}
%\item[$\minsnapshot$] Minimum duration of a snapshot in which w-correction is
%  not dominated by the Earth's curvature \vspace{-0.2cm}
%\item[$\optsnapshot$] Optimum snapshot duration (sec) \vspace{-0.2cm}
%\item[$\subint$] Sub-integration time\vspace{-0.2cm}
%\item[$\uvw$] Visibility coordinates \vspace{-0.2cm}
%\item [$\w$] Visibility $w$ coordinate \vspace{-0.2cm}
%\item[$\maxwdev$] Maximum deviation of the $w$ coordinate due to the Earth's
%  curvature \vspace{-0.2cm}
%\item[$\minwdev$] Minimum deviation of the $w$ coordinate due to the Earth's
%  curvature \vspace{-0.2cm}
%\item[$\spfiltm$] Spatial filter matrix \vspace{-0.2cm}
%\item[$\fres$] Frequency resolution \vspace{-0.2cm}
%\item[$\tres$] Time resolution (sec) \vspace{-0.2cm}
%\item[$\tressmear$]  Smearing Time resolution (sec)  \vspace{-0.2cm}
%\item[$\fracconvfuncamp$] Fraction of maximal amplitude of envelope of
%  convolution function \vspace{-0.2cm}
%\item[$\compeff$] Computational efficiency \vspace{-0.2cm}
%\item[$\fracfband$] Fraction of frequency band \vspace{-0.2cm}
%\item[$\beamsq$] Beam squint \vspace{-0.2cm}
%\item[$\fovdiam$] Angular diameter of the field of view (rad) \vspace{-0.2cm}
%\item[$\pixres$] Angular resolution of pixel size (rad) \vspace{-0.2cm}
%\item[$\beamdiam$] Angular diameter of the synthesised beam (rad)
%  \vspace{-0.2cm}
%\item[$\wavel$] Wavelength \vspace{-0.2cm}
%\item[$\maxwavel$] Maximum wavelength \vspace{-0.2cm}
%\item[$\minwavel$] Minimum wavelength \vspace{-0.2cm}
%\item[$\maxwavelsubb$] Maximum wavelength of an imaging sub-band (channels
%  over which image and pixel size are matched) \vspace{-0.2cm}
%\item[$\minwavelsubb$] Minimum wavelength of an imaging sub-band (channels
%  over which image and pixel size are matched) \vspace{-0.2cm}
%\item[$\earthrot$] Angular velocity of the Earth
%  ($7.27\times10^{-5}$\,rad/sec) \vspace{-0.2cm}
%\item[$\epsilonf$] Fraction of a uv cell that we allow uv track to move in time- or frequency- smearing-limited averaging \vspace{-0.2cm}
%\end{basedescript}

@@

 \begin{center}
 \begin{tabularx}{\textwidth}{|l|X|}
     \hline
     \bf{Reference} & \bf{Reference}\\
     \bf{Number} & \\
     \hline
   {\bf [RD01]} & SKA-TEL-SDP-0000041, F. Malan: iPython Performance Model Description\\
   {\bf [RD02]} & SKA-TEL-SDP-0000027, R. Nijboer: Pipelines Element Subsystem Design\\
   {\bf [RD03]} & SKA-TEL-SDP-0000038, R. Bolton: High Priority Science Objectives: Performance Analysis\\
   {\bf [RD04]} & SKA-TEL-SDP-0000028, A.-J. Boonstra: Ingest Pipeline\\
   {\bf [RD05]} & SKA-TEL-SDP-0000029, S. Salvini: Pipelines: Calibration\\
   {\bf [RD06]} & SKA-TEL-SDP-0000030, A. Scaife: Imaging Pipeline\\
   {\bf [RD07]} & SKA-TEL-SDP-0000031, M. Johnston-Hollitt: Science Analysis Pipeline\\
   {\bf [RD08]} & SKA-TEL-SDP-0000017, S. Wijnholds: Baseline-Dependent Averaging\\
   {\bf [RD09]} & SKA-TEL-SDP-0000058, S. Salvini: Fast Fourier Transforms\\
   {\bf [RD10]} & SKA-TEL-SDP-0000057, C. Skipper: Time and Channel Averaging\\
   {\bf [RD13]} & \url{http://www.astron.nl/casacore/trunk/casacore/doc/notes/229.html}\\
   {\bf [RD15]} & Imager.cc C++ source, CASA source code SVN revision 30821, \url{https://svn.cv.nrao.edu/svn/casa/trunk}\\ 
   {\bf [RD17]} & Parameterized deconvolution for wide-band radio synthesis imaging, Urvashi Rao Venkata, 2010, PhD thesis\\
   {\bf [RD18]} & S. Bhatnagar, T. J. (2008). Correcting direction-dependent gains in the deconvolution of radio interferometric images. A\&A, 419-429\\
   {\bf [RD22]} & Rik Jongerius, S. W. (2014). An End-to-End Computing Model for the Square Kilometre Array. IEEE Computer, volume 47, number 9\\
   {\bf [RD24]} & SDP Element Concept. Paul Alexander, Chris Broekema, Simon Ratcliffe, Rosie Bolton, Bojan Nikolic, 2013, SDP-PROP-DR-001-1 (part of SKA SDP Consortium proposal)\\
   {\bf [RD34]} & Cornwell, T. J. (2008). Multi-Scale CLEAN deconvolution of radio synthesis images. IEEE Journal of Selected Topics in Sig. Proc., 2, 793–801.\\
   {\bf [RD35]} & Conway, J. E., Cornwell, T. J., and Wilkinson, P. N. (1990). Multi-Frequency Synthesis - a New Technique in Radio Interferometric Imaging. Monthly Notices of the Royal Astronomical Society, 246, 490.\\  
   \hline
 \end{tabularx}
\end{center}

@@

\section{Purpose of the document}

@@

MS-MFS was developed by Urvashi [RD17], melding together the concepts from multi-scale clean [RD34] and multi-frequency synthesis e.g. [RD35], with the goal of improving the reconstruction of sky brightness from radio interferometric data. The explicit model used for the sky brightness is a collection of blobs of varying scale sizes and strengths, each with spectral behaviour described by a power law expanded into a Taylor series.

@@

\clearpage
\section{Overview of MSMFS}
\label{sec:overview}

@@

Before performing any search, it is necessary to decouple the pixels in Taylor term space as well as possible. In practice, the Hessian (for each scale) for the peak of the PSF in Taylor term space is of modest size and can be inverted and used. The greedy algorithm then locates the scale and location in the reference-frequency image (as convolved with the set of scales). For each minor cycle this produces a location, scale, and values for all Taylor terms. The residual images and model images are updated, and iteration proceeds. The actual algorithm is shown in Figure \ref{fig:CASA}.

@@

>>>>

\documentclass[11pt,a4paper,variablewidth]{article}

@@

\usepackage{array}
\newcolumntype{L}{>{\centering\arraybackslash}m{3cm}}

@@

\newcommand{\dg}{^\dag}
\newcommand{\X}{\vec{x}}
\newcommand{\Xd}{\vec{{x}^\dag}}
\newcommand{\B}{\vec{b}}
\newcommand{\Bd}{\vec{b^\dag}}
\newcommand{\V}{\vec{V}}
\newcommand{\Vd}{\vec{V^\dag}}
\newcommand{\A}{{\tens{A}{}}}
\newcommand{\Ad}{{\tens{A^\dag}{}}}
\newcommand{\F}{{\tens{F}{s}}}
\newcommand{\Fd}{{\tens{F^\dag}{}}}
\newcommand{\He}{{\tens{H}{}}}
\newcommand{\Sa}{{\tens{S}{}}}
\newcommand{\Sd}{{\tens{S^\dag}{}}}
\newcommand{\Sna}{\tens{{S_{\nu}}{}}}
\newcommand{\Snd}{\tens{{S_{\nu}^\dag}{}}}
\newcommand{\T}{{\tens{T}{}}}
\newcommand{\W}{{\tens{W}{}}}
\newcommand{\Wd}{{\tens{W^\dag}{}}}
\newcommand{\Pb}{{\vec{P}}}

@@

@@

 \begin{center}
 \begin{tabularx}{\textwidth}{|l|X|}
     \hline
     \bf{Reference} & \bf{Reference}\\
     \bf{Number} & \\
     \hline
   {\bf [RD01]} & SKA-TEL-SDP-0000041, F. Malan: iPython Performance Model Description\\
   {\bf [RD02]} & SKA-TEL-SDP-0000027, R. Nijboer: Pipelines Element Subsystem Design\\
   {\bf [RD03]} & SKA-TEL-SDP-0000038, R. Bolton: High Priority Science Objectives: Performance Analysis\\
   {\bf [RD04]} & SKA-TEL-SDP-0000028, A.-J. Boonstra: Ingest Pipeline\\
   {\bf [RD05]} & SKA-TEL-SDP-0000029, S. Salvini: Pipelines: Calibration\\
   {\bf [RD06]} & SKA-TEL-SDP-0000030, A. Scaife: Imaging Pipeline\\
   {\bf [RD07]} & SKA-TEL-SDP-0000031, M. Johnston-Hollitt: Science Analysis Pipeline\\
   {\bf [RD08]} & SKA-TEL-SDP-0000017, S. Wijnholds: Baseline-Dependent Averaging\\
   {\bf [RD09]} & SKA-TEL-SDP-0000058, S. Salvini: Fast Fourier Transforms\\
   {\bf [RD10]} & SKA-TEL-SDP-0000057, C. Skipper: Time and Channel Averaging\\
   {\bf [RD13]} & \url{http://www.astron.nl/casacore/trunk/casacore/doc/notes/229.html}\\
   {\bf [RD15]} & Imager.cc C++ source, CASA source code SVN revision 30821, \url{https://svn.cv.nrao.edu/svn/casa/trunk}\\ 
   {\bf [RD17]} & Parameterized deconvolution for wide-band radio synthesis imaging, Urvashi Rao Venkata, 2010, PhD thesis\\
   {\bf [RD18]} & S. Bhatnagar, T. J. (2008). Correcting direction-dependent gains in the deconvolution of radio interferometric images. A\&A, 419-429\\
   {\bf [RD22]} & Rik Jongerius, S. W. (2014). An End-to-End Computing Model for the Square Kilometre Array. IEEE Computer, volume 47, number 9\\
   {\bf [RD24]} & SDP Element Concept. Paul Alexander, Chris Broekema, Simon Ratcliffe, Rosie Bolton, Bojan Nikolic, 2013, SDP-PROP-DR-001-1 (part of SKA SDP Consortium proposal)\\
   {\bf [RD34]} & Cornwell, T. J. (2008). Multi-Scale CLEAN deconvolution of radio synthesis images. IEEE Journal of Selected Topics in Sig. Proc., 2, 793–801.\\
   {\bf [RD35]} & Conway, J. E., Cornwell, T. J., and Wilkinson, P. N. (1990). Multi-Frequency Synthesis - a New Technique in Radio Interferometric Imaging. Monthly Notices of the Royal Astronomical Society, 246, 490.\\
   {\bf [RD36]} & http://www.aanda.org/component/citedby/?task=crossref\&doi=10.1051/0004-6361/201117104 \\  
   {\bf [RD37]} & Junklewitz, H., Bell, M. R., and Ensslin, T. (2015). A new approach to multifrequency synthesis in radio interferometry. Astronomy \& Astrophysics, 581, A59. http://doi.org/10.1051/0004-6361/201423465 \\
   \hline
 \end{tabularx}
\end{center}

@@

\clearpage
\section{Purpose of the document}

@@

The imaging performance of a radio synthesis telescope is related to the uv coverage. The uv coverage is dependent on the observing wavelength so diversity in uv can be obtained by observing over a range of frequencies [RD35]. This Multi-Frequency Synthesis (MFS) only works for radio sources in which broadband emission is present, giving a way to relate measurements at different frequencies. There have been multiple attempts at an MFS algorithm (e.g. [RD35]). MFS is usually most important for sources having extended emission. Hence multi-scale algorithms are favoured.

The Multi-Scale, Multi-Frequency Synthesis (MS-MFS) algorithm was developed by Urvashi [RD17], melding together the concepts from multi-scale clean [RD34] and multi-frequency synthesis e.g. [RD35], with the goal of improving the reconstruction of sky brightness from radio interferometric data. The explicit model used for the sky brightness is a collection of blobs of varying scale sizes and strengths, each with spectral behaviour described by a power law expanded into a Taylor series. It has been used in JVLA, ALMA, and ATCA observations [RD36].

@@

\section{Is MFS needed?}
\label{sec:mfsneeded}

There is an L1 requirement, $SKA1-SYS\_REQ-2324$ specifying that MSMFS must be used (see Table\ref{tab:L1}).

\begin{table}[htp]
\begin{tabular}{|c|L|c|c|c|}
\hline
ID & Requirement & Status & Parent Requirement & Verification \\
\hline
$SKA1-SYS\_REQ-2324$ &Multi-frequency synthesis imaging. All imaging shall construct and make use of frequency dependent image models over the entire observed bandwidth. & Accepted & Reference [2] & Demonstration \\
\hline 
\end{tabular}	
\caption{Relevant L1 requirement}\label{tab:L1}	
\end{table}

This is reflected in the L2 requirements shown in Table\ref{tab:L2}.
\begin{table}[htp]
\begin{tabular}{|L|L|L|L|L|}
\hline
\multicolumn{2}{|c|}{Function}&\multicolumn{3}{c|}{Satisfies requirement}\\
\hline
Number & 	Name & 	Number	& Name	& Description \\
\hline
F1.9 & Calibrate and Image & $SDP\_REQ-390$	&Multi-frequency synthesis
  imaging	& The SDP shall construct and make
  use of frequency-dependent image models over the entire observed bandwidth. \\
\hline
F1.9 & Calibrate and Image & $SDP\_REQ-676$	& Continuum imaging deconvolution
  scales &	While performing continuum
  imaging the SDP shall deconvolve using a maximum number of 10 (TBC) scales.\\
\hline
F1.9 & Calibrate and Image & $SDP\_REQ-677$	& Continuum imaging deconvolution
  Taylor terms &	While performing continuum
  imaging the SDP shall deconvolve using a maximum number of 5 (TBC) Taylor
  terms.\\
\hline
F1.9 & Calibrate and Image & $SDP\_REQ-678$	& Continuum imaging deconvolution
  by sub-band	& While performing continuum
  imaging the SDP shall deconvolve by sub-band (TBC).\\
\hline 
\end{tabular}
\caption{Relevant L2 requirements}	\label{tab:L2}
\end{table}

The last three L2 requirements, $SDP\_REQ-676$, $SDP\_REQ-677$, $SDP\_REQ-678$, are an attempt to give some constraints to $SKA1-SYS\_REQ-2324$. However, they beg the same question as the original L1 requirement: is MSMFS needed to meet the scientific goals? If so, how many scales, Taylor terms, and sub-bands are required? In addition, there is also a potentially missing requirement concerning the connection between the models for different sub-bands - is there allowed to be a seam in the continuum-subtracted spectral line cube at the transitions between sub-bands?

Finally we could ask is there a superior algorithm already published? (see e.g. [RD37]).

Some caution is advised in advocating that any form of MFS is not needed. Clearly ALMA and JVLA, which also have relatively good uv coverage, have benefited from the use of MSMFS. Furthermore, although the uv-coverage of SKA1-LOW and SKA1-MID is excellent after a long observation, there will inevitably be pressure to observe with shorter observation times. We should remember than the VLA was supposed to take 8 hours to image every field. With the advent of CLEAN and self-calibration, it was known even in the early eighties that 400 sources could be imaged in a day.

There is no mechanism to address these questions beyond simulations. Arguing from uv coverage metrics such as the UVGAP only helps in a relative sense.

\section{Overview of MSMFS}
\label{sec:overview}

@@

@@

The visibility function corresponding to \ref{eqn:msmf_model} is:
\begin{eqnarray}
\vec{V}^{\rm obs}_{\nu} &=& [\Sna][\F]\vec{I}^{\rm m}_{\nu}
= \sum_{t=0}^{\Nt } \sum_{s=0}^{\Ns } \wnt [\Sna][\T_s][\F] \vec{I}^{\rm sky}_{s\atop t}
\label{eqn:msmfs_meqn}
\end{eqnarray}

$\vec{V}^{\rm obs}_{\nu}$ is a vector of $n\times 1$ visibilities
measured at frequency $\nu$.
$w_{\nu}$ are Taylor-weights (shown in Eqn.\ref{Eq:msmf_model}).
$[\Sna]$ is an $n\times m$ projection operator that represents the 
spatial-frequency sampling function for frequency $\nu$. 
The image-domain convolution of model $\vec{I}^{\rm sky}_{s\atop t}$ with
 $\I^{\rm shp}_s$ is written as a spatial-frequency-domain multiplication.
$ \vec{I}^{\rm shp}_s \star \vec{I}^{\rm sky}_{s\atop t} = [\Fd][\T_s][\F] \vec{I}^{\rm sky}_{s\atop t}$
where $[\T_s]_{m\times m} = diag([\F] \vec{I}^{\rm shp}_s)$ is a spatial-frequency
taper function.  All images are vectors of shape $m\times 1$.

The normal equations 
for the system described in Eqn. \ref{Eq:msmfs_meqn}
can be written in block matrix form, with each block-row 
(for scale size $s$, and Taylor term $t$) given by
\begin{eqnarray}
\label{Eq:msmfs_neqn_1}
\sum_{p=0}^{\Ns -1}\sum_{q=0}^{\Nt -1} \left[\He_{{s,p}\atop{t,q}}\right] \vec{I}^{\rm sky}_{{p}\atop{q}} &=& \vec{I}^{\rm dirty}_{{s}\atop{t}}~~~  \forall~ s \in [0,\Ns -1], t\in[0,\Nt -1]
\end{eqnarray}
Here, each $\left[\He_{{s,p}\atop{t,q}} \right]$ is an $m\times m$ block of the 
Hessian matrix, and $\vec{I}^{\rm dirty}_{{s}\atop{t}}$ is one of $\Ns  \Nt $ dirty images.
\begin{eqnarray}
\left[\He_{{s,p}\atop{t,q}} \right] &=& \left[\A_{{s}\atop{t}}^{\dag}\right][\Wim] \left[\A_{{p}\atop{q}}\right] \\
 &=&    [\Fd \T_s  \Sd \Wntd ]  [\Wim]  [\W^{\rm mfs}_q \Sa \T_p \F] \\
 &=&    [\Fd \T_s \F] [\Fd  \Sd \Wntd  \Wim  \W^{\rm mfs}_q \Sa  \F] [\Fd \T_p \F]\\
% &=&    [\Fd \T_s \F] [\He^{\rm mfs}_{t,q}] [\Fd \T_p \F]
 &=& [\Fd \T_s \F] \left\{  \sum_{\nu} \wntq [\Fd\Snd\Wimn\Sna\F] \right\} [\Fd \T_p \F] \\
 &=& [\Fd \T_s \F] \left\{  \sum_{\nu} \wntq [\He_{\nu}\} \right\} [\Fd \T_p \F] 
\end{eqnarray}

$[\Wim]$ is a diagonal matrix of data-weights (and imaging-weights) and  
$[\W^{\rm mfs}_t]$ is a diagonal matrix containing Taylor-weights $\wnt$.
$[\He_{\nu}] =  [\Fd\Snd\Wimn\Sna\F]$ is the Hessian matrix formed using only one
 frequency channel, and
is a convolution operator containing a shifted version of the single-frequency 
point-spread-function $\I^{\rm psf}_{\nu} = diag[\Fd \Sd \Wimn \Sa]$ in each row 
(see appendix \ref{App:normal} for details).
$[\Fd \T_s\F]$ and $[\Fd\T_p\F]$ are also convolution operators 
with $\I^{\rm shp}_s$ and $\I^{\rm shp}_p$ as their kernels. 
The process of convolution is associative and commutative, and 
therefore, $\left[\He_{{s,p}\atop{t,q}}\right]$  is also a convolution operator 
whose kernel is given by 
\begin{equation}
\label{Eq:msmfs_neqn_2.5}
\vec{I}^{\rm psf}_{{s,p}\atop{t,q}} = \I^{\rm shp}_s  \star \left\{ \sum_{\nu} \wntq \vec{I}^{\rm psf}_{\nu} \right\} \star \I^{\rm shp}_p 
%&=& \I^{\rm shp}_s \star I^{\rm psf}_{t,q} \star \I^{\rm shp}_p
\end{equation}

\noindent The dirty images on the RHS of Eqn.\ref{Eq:msmfs_neqn_1} can be written as follows.
\begin{eqnarray}
\label{Eq:msmfs_neqn_3}
\vec{I}^{\rm dirty}_{{s}\atop{t}}  &=&  \left[\A_{{s}\atop{t}}^{\dag}\right][\Wim]  \vec{V}^{\rm obs}\\
&=& [\Fd \T_s \Sd\Wntd\Wim] \vec{V}^{\rm obs} \\
&=& [\Fd \T_s \F][\Fd\Sd\Wntd\Wim] \vec{V}^{\rm obs} \\
&=& [\Fd \T_s \F] \left\{  \sum_{\nu} \wnt [\Fd\Snd\Wimn] \vec{V}_{\nu}^{\rm obs}  \right\} \\
&=& \I^{\rm shp}_s \star \left\{\sum_{\nu} \wnt \vec{I}^{\rm dirty}_{\nu}  \right\} 
%&=& [\Fd \T_s \F] \vec{I}^{\rm dirty}_t = \I^{\rm shp}_s \star \vec{I}^{\rm dirty}_t\\
\label{Eq:msmfs_neqn_3a}
%\vec{I}^{\rm dirty}_{t}  &=& [\Fd\Sd\Wntd\Wim] \vec{V}^{corr}\\
% &=& \sum_{\nu} \wnt [\Fd\Snd\Wimn] \vec{V}_{\nu}^{corr} = \sum_{\nu} \wnt \vec{I}^{\rm dirty}_{\nu}\\
%\label{Eq:specPSF}
%\vec{I}^{\rm psf}_{t,q}&=& \sum_{\nu} \wntq [\Fd\Snd\Wimn] \vec{1} = \sum_{\nu} \wntq \vec{I}^{\rm psf}_{\nu}
\end{eqnarray}
where ${\I}_{\nu}^{\rm dirty} = [\Fd \Snd \Wimn] \vec{V}^{\rm obs}_{\nu} $ is the dirty
image formed by direct Fourier inversion of weighted visibilities from one frequency channel.

When all scales and Taylor terms are combined, 
the full Hessian matrix contains
$\Nt  \Ns  \times \Nt  \Ns $ blocks each of size $m\times m$, 
and $\Nt $ Taylor coefficient images each of size $m\times 1$, 
for all $\Ns $ spatial scales.

The normal equations in block matrix form for the example 
in Eqn.\ref{meqn_msmfs_math} for $\Nt =3, \Ns =2$ is shown
in Eqn.\ref{Eq:msmfs_neqn_matrix}.
The Hessian matrix consists of $\Ns \times \Ns =2\times 2$ blocks 
(the four quandrants of the matrix), each for one pair of spatial scale $s,p$ (the upper indices).
Within each quadrant, the $\Nt \times \Nt  = 3\times 3$
matrices correspond to various pairs of $t,q$ (Taylor coefficient indices; the lower indices).
This layout shows how the multi-scale and multi-frequency aspects of
this imaging problem are combined and illustrates the 
dependencies between the spatial and spectral basis functions.
